Supplement 2

### 2020-11-19 paper/paper.Rmd
edits before preprint.

17 November 2020
\vspace{5mm}

@@

1: Griffith University, Menzies Health Institute Queensland, The Hopkins Centre, Brisbane, Australia.\hfill \break
2: Griffith University, School of Allied Health Sciences, Brisbane, Australia.\hfill \break
3: University of New South Wales, Department of Statistics, School of Mathematics and Statistics, Sydney, Australia.\hfill \break
4: Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia.\hfill \break
5: Australian Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS), Melbourne, Australia.

@@

Missing data are often unavoidable. The reason values go missing, along with decisions made of how missing data are handled (deleted or imputed), can have a profound effect on the validity, accuracy and usability of a study's results. In this article, we aimed to: estimate the proportion of studies in football research that included a missing data statement, highlight several practices to avoid in relation to missing data, and provide recommendations for exploring, visualising and reporting missingness. Football related articles, published in 2019 were studied. A survey of `r n_screened` articles, sampled at random, was conducted to determine whether a missing data statement was included. As expected, the proportion of studies in football research that included a missing data statement was low, at only `r binom_ci_miss_reported$mean_pct` (95% CI: `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`); suggesting that missingness is seldom considered by researchers. It is suspected that this result is consistent with practice in the wider sports science literature. We recommend researchers describe the number and percentage of missing values, including when there are no missing values. Exploratory analysis should be conducted to explore missing values, and visualisations describing missingness overall should be provided in the paper, or at least supplementary materials. Missing values should almost always be imputed, and imputation methods should be explored to ensure they are appropriately representative.

@@

For clarity, we describe an example of missingness. An Australian Football data collector was taking lunch during the third quarter, so no data was recorded for this time period. This would be considered missing data. Compare this to the game being cancelled due to a pandemic. There is no missing data because there is no data to be observed. Similarly, if a study screens out individuals who do not meet inclusion criteria, this is not a missing value, because there is no intent for these values to be observed or measured.

@@

It is imperative that missing data, or lack thereof, are reported [@Schafer2002]. While there has been significant interest in modelling missing data [@Tierney2015; @Barnett2017], the exploration and reporting of missings have received less attention [@Schafer2002]. This article aimed to: (1) estimate the proportion of articles that report missing data in football research; (2) highlight several practices that should be avoided; and (3) provide recommendations for exploring and reporting missingness.

@@

To estimate the proportion of articles that report missing data in football research, we conducted a systematic search [@Moher2009]. Articles on football topics, or involving footballers, published in 2019 were studied. Three major sports medicine databases (SPORTDiscus, Embase and Cinahl) were searched, using key terms and search limits (Supplement 2). Included articles were: (a) written in English; (b) had an accessible full-text; and (c) included quantitative data. The PRISMA figure in Supplement 2 summarises the search process, and further details of the search can also be found in Supplement 2.

@@

Our intention was to survey, at random, 10% of the total articles found after title and abstract screening. We were primarily interested in whether missing data was reported or acknowledged; and if not, whether a dataset was shared. Where applicable, information relating to how missing values were treated was also extracted (see Supplement 2). Two authors independently extracted the data (`r md_reported` agreement). Disagreements were resolved by author DNB. Results are reported as the proportion and 95% confidence interval (CI).

@@

Of the 199 articles screened, `r n_screened` met the inclusion criteria, representing \~10% of the articles found (see Supplement 2, Figure 1). The proportion of articles that reported missing data was `r binom_ci_miss_reported$mean_pct` (95% CI: `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`). Of the articles that did not mention missing data (`r n_no_mention_missings`/`r n_screened`), `r binom_ci_data_shared$mean_pct` (95% CI: `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`) shared their dataset, making
it difficult to determine whether there was any missingness. While the
absence of a missing data statement does not necessarily mean missing
data are not being reported, it does suggest that this aspect of the
analysis process receives little consideration.

@@

Our survey of `r n_screened` articles published on football related topics found that only one in nine papers (`r binom_ci_miss_reported$mean_pct`) reported missing data. It is reasonable to assume that this result is representative of current practice in the wider sports science literature. The low proportion of papers reporting missingness could be explained by a lack of awareness and education [@Sainani2020call], similar to other sciences [@Peugh2004; @Jelivcic2009]. Below we discuss several practices to avoid in relation to missingness, and provide recommendations for exploring, visualising and reporting missing data.

@@

*Unreported missing values.* While not always the case, even if
not reported, missing data can be obvious. For example, a study examined the
relationship between match performance indicators and outcome, in
Australian Football between the 2001 and 2016 [@Young2019], where at least one of the
91 performance indicators ('meters gained') included in the analysis was not available over the entire 2001-2016 period [@Young2019mentionofmissing]. This was not noted anywhere in the paper, or in the supplementary
materials. It is unknown whether imputation was undertaken. Without
these details, it is difficult for other analysts and researchers to use
and/or extend the ideas presented in the paper, or reproduce the
analysis. It is also difficult for a reader to evaluate how the
missingness could bias the results. Authors should include a missing data statement,
irrespective of whether data are missing, or not.

*Mean imputation.* The default of most statistical software for
handling missing data in modelling is listwise deletion. Listwise
deletion removes entire rows that contain missing observations from the
analysis. At best, listwise deletion reduces the power in analysis by
reducing the sample size. At worst, it introduces bias. For example, if
injured participants are removed from a study on injury prevention.
Imputing values removes the need to conduct listwise deletion. However,
it is critical how these values are imputed, and that the imputation
method is documented. Authors sometimes choose to replace missing 
observations with mean values, such as a participant or group average [@Sampson2018; @Cresswell2009]. 
While imputing the mean preserves the study sample size (and point estimates), it also reduces
the variance, can alter the relationship between the variable with
missing observations and other variables, and can bias (underestimate)
standard errors [@Scheffer2002]. Smaller standard errors typically reduce *p*-values, which may lead to
incorrect inference. Mean imputation should be avoided. An imputation
strategy that incorporates information from other related variables in
the data---such as linear regression, or K nearest neighbours---should
be considered.

@@

Broadly, there are two types of missing values in data: implicit and explicit missings. Explicit values are missing, but recorded; whereas information in the data. For example, in Figure 1, player Koenen has missing values for quarters two and four. Sometimes values like these can be logically imputed, as it might be known that these values are recorded as 0 in this format, rather than NA. Other places missing data can arise include: through joins when merging data without corresponding values, surveys, an inability to collect a biological sample (e.g., venous blood), equipment malfunction, failure, or not being worn.

@@

```{r implicit-missings, fig.cap = "Two tables demonstrating explicit and implicit missing. The first table shows the number of goals scored for a player in a given quarter of an AFL match with the first column showing the player name, the second the quarter they played, and the third the goals they scored. Note that Player, 'Koenen' has no entries for Quarter 2 and 4. The second table shows the same information from the first table pivoted, with each row being a player and the number of goals they scored in each quarter, with each quarter being a column. We notice that in the second form of the data, we can clearly see that Koenen has missing values. These types of missing values have a name, implicit missing values. The first table has implicit missing values, meaning they are implied, and the second table has those implicit missing values explicitly expressed."}

@@

```{r missing-overview,  fig.cap = "Overviews of missing values in simulated injury data from Australian Football matches played in August, 2020. The dataset contains observed player id, team, and match venue data, in addition to simulated data on injury incidence, and number of sprints per match, and total sprint distance per match. Panel A shows the location of missing (black) and complete (grey) cases, and the percentage of missing data in each variable and overall. Panel B shows the percentage of missings for each team, in each variable. Panel C shows the relationship between two variables in the dataset, including missing values (red) which are shifted, so that they can be visualised on the same axes as the regular values. Panel D shows the number of missing values in each variable", fig.height = 7.75}

@@

Missing values can occur in the both the predictors (the "independent variables") and the response variable (the "dependent variables"). Care should be taken when imputing the response variables, and in general they should not be imputed with a single imputation, as this will bias the subsequent analysis. Methods that provide multiple imputed values for a response variable, such as multiple imputation, or using a Bayesian framework that generates a distribution of posterior values, can be appropriate for imputing the response variable as they describe the uncertainty---see Chapter 15 in [@McElreath2020]. Missing values in the outcome can be dealt with specially for Randomised Control Trials [@Jakobsen2017; @Wood2004].

Much in the way of there is no single 'best' statistical method, there is no perfect, one-size-fits-all approach for imputing data. The goal is to generate similar values that might have been otherwise recorded. Sometimes this means using a neighborhood approach of finding similar values. Or it could mean predicting responses using linear or tree models. Other times the most likely value might have been 0 or the last or first value carried forward. For detailed descriptions, and a summary of these methods, we suggest [@VanBuuren2012; @Schafer2002; @Cheema2014].

@@

The current state of reporting missing data in football research has room for improvement. In this section we discuss recommended practices for exploring and reporting missing data. Our intention is not to provide a 'cookbook' style approach to missing data, but rather, broad recommendations to help authors when writing the methods and results sections of a study, and assist researchers when evaluating a study during the peer-review process.

>>>>

20 November 2020
\vspace{5mm}

@@

1: Griffith University, Menzies Health Institute Queensland (MHIQ), The Hopkins Centre, Brisbane, Australia. 2: Griffith University, School of Allied Health Sciences, Brisbane, Australia. 3: University of New South Wales, Department of Statistics, School of Mathematics and Statistics, Sydney, Australia. 4: Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia. 5: Australian Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS), Melbourne, Australia.

@@

Missing data are often unavoidable. The reason values go missing, along with decisions made of how missing data are handled (deleted or imputed), can have a profound effect on the validity and accuracy of study results. In this article, we aimed to: estimate the proportion of studies in football research that included a missing data statement, highlight several practices to avoid in relation to missing data, and provide recommendations for exploring, visualising and reporting missingness. Football related articles, published in 2019 were studied. A survey of `r n_screened` articles, sampled at random, was conducted to determine whether a missing data statement was included. As expected, the proportion of studies in football research that included a missing data statement was low, at only `r binom_ci_miss_reported$mean_pct` (95% CI: `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`); suggesting that missingness is seldom considered by researchers. We recommend that researchers describe the number and percentage of missing values, including when there are no missing values. Exploratory analysis should be conducted to explore missing values, and visualisations describing missingness overall should be provided in the paper, or at least supplementary materials. Missing values should almost always be imputed, and imputation methods should be explored to ensure they are appropriately representative. Researchers should consider these recommendations, and pay greater attention to missing data and its influence on research results.

@@

For clarity, we describe an example of missingness. An Australian Football data collector was taking lunch during the third quarter, so no data was recorded for this time period. This would be considered missing data. Compare this to the game being cancelled due to a pandemic. There is no missing data because there is no data to be observed. Similarly, if a study screens out individuals who do not meet inclusion criteria, these are not missing values, because there is no intent for these values to be observed or measured.

@@

It is imperative that missing data, or lack thereof, are reported [@Schafer2002]. While there has been significant interest in modelling missing data [@Tierney2015; @Barnett2017], the exploration and reporting of missings have received less attention [@Schafer2002]. The current article aimed to: (1) estimate the proportion of studies that report missing data in football research; (2) highlight several practices that should be avoided; and (3) provide recommendations for exploring and reporting missingness.

@@

To estimate the proportion of articles that report missing data in football research, we conducted a systematic search [@Moher2009]. Articles on football topics, or involving footballers, published in 2019 were studied. Three major sports medicine databases (SPORTDiscus, Embase and Cinahl) were searched, using key terms and search limits (Supplement 2). Included articles were: (a) written in English; (b) had an accessible full-text; and (c) included quantitative data. The PRISMA figure in Supplement 2 summarises the search process.

@@

Our intention was to survey, at random, 10% of the total articles found after title and abstract screening. We were primarily interested in whether missing data was reported or acknowledged; and if not, whether a dataset was shared. Where applicable, information relating to how missing values were treated was also extracted (Supplement 2). Two authors independently extracted the data (`r md_reported` agreement). Disagreements were resolved by author DNB. Results are reported as the proportion and 95% confidence interval (CI).

@@

Of the 199 articles screened, `r n_screened` met the inclusion criteria, representing \~10% of the articles found (see Supplement 2, Figure 1). The proportion of articles that reported missing data was `r binom_ci_miss_reported$mean_pct` (95% CI: `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`). Of the articles that did not mention missing data (`r n_no_mention_missings`/`r n_screened`), only `r binom_ci_data_shared$mean_pct` (95% CI: `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`) shared their dataset, making
it difficult to determine whether there was any missingness. While the
absence of a missing data statement does not necessarily mean missing
data are not being reported, it does suggest that this aspect of the
analysis process receives little consideration.

@@

Our survey of `r n_screened` articles published on football related topics found that only one in nine papers (`r binom_ci_miss_reported$mean_pct`) reported missing data. It is reasonable to assume that this result is representative of current practice in the wider sports science literature. The low proportion of papers reporting missingness might be explained by a lack of awareness and education [@Sainani2020call], similar to other sciences [@Peugh2004; @Jelivcic2009]. Below we discuss several practices to avoid in relation to missingness, and provide recommendations for exploring, visualising and reporting missing data.

@@

*Unreported missing values.* While not always the case, even if
not reported, missing data can be obvious. For example, a study examined the
relationship between match performance indicators and outcome, in
Australian Football between the 2001 and 2016 [@Young2019], where at least one of the
91 performance indicators ('meters gained') included in the analysis was not available over the entire 2001-2016 period [@Young2019mentionofmissing]. This was not noted anywhere in the paper, or in the supplementary
materials. It is unknown whether imputation was used. Without
these details, it is difficult for other analysts and researchers to use
and/or extend the ideas presented in the paper, or reproduce the
analysis. It is also difficult for a reader to evaluate how the
missingness could bias the results. Authors should include a missing data statement,
irrespective of whether data are missing, or not.

*Mean imputation.* The default of most statistical software for
handling missing data in modelling is listwise deletion. Listwise
deletion removes entire rows that contain missing observations from the
analysis. At best, listwise deletion reduces the power in the analysis, by
reducing the sample size. At worst, it introduces bias. For example, if
injured participants are removed from a study on injury prevention.
Imputing values removes the need to conduct listwise deletion. However,
it is critical how these values are imputed, and that the imputation
method is documented. Authors sometimes choose to replace missing 
observations with mean values, such as a participant or group average [@Sampson2018; @Cresswell2009]. 
While imputing the mean preserves the study sample size (and point estimates), it also reduces
the variance, can alter the relationship between the variable with
missing observations and other variables in the dataset, and can bias (underestimate)
standard errors [@Scheffer2002]. Smaller standard errors typically reduce *p*-values, which may lead to
incorrect inference. Mean imputation should be avoided. An imputation
strategy that incorporates information from other related variables in
the data---such as linear regression, or K nearest neighbours---should
be considered.

@@

Broadly, there are two types of missing values in data: implicit and explicit missings. Explicit values are missing, but recorded; whereas implicit missings are not recorded in the dataset but implied. For example, in Figure 1, player Koenen has missing values for quarters two and four. Sometimes values like these can be logically imputed, as it might be known that these values should have been recorded as 0, rather than NA. Other places missing data can arise include: through joins when merging data without corresponding values, on surveys, an inability to collect a biological sample (e.g., venous blood), or due to equipment malfunction, failure, or not being worn.

@@

```{r implicit-missings, fig.cap = "Two tables demonstrating explicit and implicit missings. The first table shows the number of goals scored for a player in a given quarter of an AFL match with the first column showing the player name, the second column the quarter they played, and the third column the goals they scored. Note that Player, 'Koenen' has no entries for Quarters 2 and 4. The second table shows the same information from the first table pivoted, with each row being a player and the number of goals they scored in each quarter, with each quarter being a column. We notice that in the second form of the data, we can clearly see that Koenen has missing values. These types of missing values have a name, implicit missing values. The first table has implicit missing values, meaning they are implied, and the second table has those implicit missing values explicitly expressed."}

@@

```{r missing-overview,  fig.cap = "Overviews of missing values in simulated injury data from Australian Football matches played in August, 2020. The dataset contains observed player id, team, and match venue data, in addition to simulated data on injury incidence, and number of sprints per match, and total sprint distance per match. Panel A shows the location of missing and present cases, and the percentage of missing data in each variable (and overall). Panel B shows the percentage of missings for each team, in each variable. Panel C shows the relationship between two variables in the dataset, including missing values (red) which are shifted, so that they can be visualised on the same axes as the regular values. Panel D shows the number of missing values in each variable", fig.height = 7.75}

@@

Missing values can occur in the both the predictors (the "independent variables") and the response variable (the "dependent variables"). Care should be taken when imputing the response variable, and in general they should not be imputed with a single imputation, as this will bias the subsequent analysis. Methods that provide multiple imputed values for a response variable, such as multiple imputation, or using a Bayesian framework that generates a distribution of posterior values, can be appropriate for imputing outcome variables as they describe the uncertainty (see Chapter 15 in [@McElreath2020]). Missing values in the outcome can be dealt with specially for Randomised Control Trials [@Jakobsen2017; @Wood2004].

Much in the way of there is no single 'best' statistical method, there is no perfect, one-size-fits-all approach for imputing data. The goal is to generate similar values that might have been otherwise recorded. Sometimes this means using a neighborhood approach of finding similar values. Or it could mean predicting responses using linear or tree models. Other times the most likely value might have been 0 or the last (or first) value carried forward. For detailed descriptions, and a summary of these methods, we suggest [@VanBuuren2012; @Schafer2002; @Cheema2014].

@@

The current state of reporting missing data in football research has room for improvement. In this section we recommend practices for exploring and reporting missing data. Our intention is not to provide a 'cookbook' style approach to missing data, but rather, broad recommendations to help authors when writing the methods and results sections of a study, and assist researchers when evaluating a study during the peer-review process.

### 2020-11-19 paper/supplement-1.Rmd
edits before preprint.

---
title: Supplement 1
output: pdf_document
bibliography: supp1.bib
---

@@

Data MAR could arise if players' body mass measurements were taken with a
weighing scale that produced more missing values when placed on soft carpet,
rather than hard tiles, in the team change room. If surface type was recorded,
data would be MAR, because the missingness is related the surface, but not
related to the value of body mass itself [@VanBuuren2012].

Data MNAR is when the missings are related to something unobservable or unobserved. The previous example of data MAR could be data MNAR if the surface type was not recorded in the dataset---it would then be unobserved. Data MNAR also arises when missing values are related to the values missing. This could take place if, for example, the scales used only measure up to 110 Kg, and then produce missing values. This is a systematic issue---it happens each time, and means that heavier players are not recorded.

>>>>

---
title: Supplement 1
output: pdf_document
bibliography: supp1.bib
csl: sport-medicine.csl
---

@@

Data MAR could arise if players' body mass measurements were taken with a scale 
that produced more missing values when placed on soft carpet,
rather than hard tiles, in the team change room. If surface type was recorded,
data would be MAR, because the missingness is related the surface, but not
related to the value of body mass itself [@VanBuuren2012].

Data MNAR is when the missings are related to something unobservable or unobserved. The previous example of data MAR could be data MNAR if the surface type was not recorded in the dataset---it would then be unobserved. Data MNAR also arises when missing values are related to the values missing. This could take place if, for example, the scales used only measure up to 110 kg, and then produce missing values. This is a systematic issue---it happens each time, and means that heavier players are not recorded.

### 2020-11-19 paper/supplement-2.Rmd
edits before preprint.

---
title: "Supplement 2"
output:
  bookdown::html_document2:
    keep_md: true
  bookdown::pdf_document2:
    toc: false
bibliography: "supp2.bib"
---

@@

# Systematic search details

The systematic search was performed in accordance with the reporting
requirements of the PRISMA statement [@Moher2009]. The search strategy is
described below. The search protocol was not registered, and no funding
was provided to support the undertaking of the search.

@@

This review aimed to estimate the proportion of articles that reported 
missing data in football research.

@@

Football related articles published in 2019 were studied. Three major
sports medicine databases were searched on May 20, 2020, using key terms and limits. 
The databases were: SPORTDiscus, Embase and Cinahl. A
list of the search terms has been provided in
Table \@ref(tab:search-table-print) below.

```{r search-table}
table_terms <- tibble(
  SPORTDiscus = c("exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching",
                  "AND",
    "football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football"),
  Embase = c("(exercise AND science OR (sports AND science) OR physiology OR biomechanics OR nutrition OR training OR testing OR (sports AND medicine) OR 'performance analysis' OR (('performance'/exp OR performance) AND ('analysis'/exp OR analysis)) OR 'performance'/exp OR performance OR 'sport psychology'/exp OR 'sport psychology' OR (('sport'/exp OR sport) AND ('psychology'/exp OR psychology)) OR 'coaching'/exp OR coaching) AND [2019-2019]/py",
             "AND",
              "(football OR soccer OR (indoor AND soccer) OR futsal OR rugby OR (rugby AND league) OR (rugby AND union) OR (gaelic AND football) OR 'australian rules footbal' OR (('australian'/exp OR australian) AND rules AND footbal) OR afl OR 'american football'/exp OR 'american football' OR (('american'/exp OR american) AND ('football'/exp OR football)) OR 'gridiron football' OR (gridiron AND ('football'/exp OR football)) OR 'touch football' OR (('touch'/exp OR touch) AND ('football'/exp OR football))) AND [2019-2019]/py"),
  Cinahl = c("exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching",
             "AND",
              "football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football")

)
```

```{r search-table-print}
table_terms %>% 
  kbl(caption = "Key search terms and complete search statements for the databases, SPORTDiscuss, Embase, and Cinahl",
      booktabs = TRUE) %>% 
  kable_styling("striped",
                full_width = TRUE)
```

Included articles were: (a) published in 2019; (b) in English; 
(c) had an accessible full-text; and (d) included quantitative data. 
Studies containing original data, including case studies (with time-series data) 
and n=1 trials, were considered. The search results were imported into EndNote (version 9.3.2).
Duplicates were removed, and titles and abstracts were screened against
the inclusion criteria. Of the articles identified for full-text
screening, 10% were sampled, at random, for survey. Because \~30% of the
papers from the initial random sample were not eligible for inclusion
(e.g., full-texts not in English, conference abstracts, qualitative
studies), an additional 5% random sample was taken. It was expected that 
30% of the additional 5% random sample would also be ineligible. With this loss,
it was expected that 10% of the articles that remained after title and abstract 
screening would undergo survey. 
Figure \@ref(fig:prisma) summarises the search process.

```{r prisma, fig.cap = "Flowchart of the article search and inclusion for the systematic search. The full-text of 137 articles was reviewed.", warning = FALSE, fig.height = 10, fig.width = 8}
knitr::include_graphics("figures/prisma-figure.pdf")
```

Articles were surveyed to determine: (a) whether a missing data
statement was included, or if missing data were acknowledged in a
figure/table; where data were missing, whether (b) data were considered
MCAR, MAR or MNAR; (c) imputation was used; (d) the method used for imputation; 
and (e) whether articles that did not include a missing data
statement shared their data. Two authors independently extracted the
data. Discrepancies were resolved by DNB. Findings are
reported as the proportion and 95% confidence interval (CI). Confidence
intervals were calculated using the Clopper-Pearson method for the
binomial distribution [@Clopper1934] via the *binom* package [@binom] in R [@rcore].

@@

The search returned a total of 3,863 articles. Of the 199 articles screened, `r n_screened` met the inclusion criteria, representing \~10% of the articles found (Figure \@ref(fig:prisma)). The proportion of articles that reported missing data was `r binom_ci_miss_reported$mean_pct` (95% CI; `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`). Of the articles that did not mention missing data, `r n_no_mention_missings` / `r n_screened`, `r binom_ci_data_shared$mean_pct` (95% CI; `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`) shared their dataset, making it difficult to determine whether there was any missingness. While the absence of a missing data statement does not necessarily mean missing data are not being reported, it does suggest that this aspect of the analysis process receives little consideration.

There were `r nrow(gs_football_missings_stated)` studies that provided a statement about missing data. A breakdown of the extracted information from these studies is provided in Table \@ref(tab:extracted-info).

@@

A survey of `r n_screened` articles published in 2019 on football related topics
showed that only one in nine reported missing data, suggesting
that missing data are rarely considered by sport science researchers.

>>>>

---
title: "Supplement 2"
output:
  pdf_document:
    toc: false
  html_document:
    keep_md: true
bibliography: "supp2.bib"
csl: sport-medicine.csl
---

@@

# Systematic Search Details

The systematic search was performed in accordance with the reporting
requirements of the PRISMA statement [@Moher2009]. The search strategy is
described below. The search protocol was not registered, and no funding
was provided to support the undertaking of this work.

@@

We aimed to estimate the proportion of articles that reported 
missing data in football research.

@@

Football related articles published in 2019 were studied. Three major
sports medicine databases were searched on May 20, 2020, using key terms and limits. 
The databases were: SPORTDiscus, Embase and Cinahl. The search terms were: exercise science 
OR sports science OR physiology OR biomechanics OR nutrition OR training 
OR testing OR sports medicine OR performance analysis OR performance 
OR sport psychology OR coaching AND football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football.

Included articles were: (a) published in 2019; (b) in English; 
(c) had an accessible full-text; and (d) included quantitative data. 
Studies containing original data, including case studies (with time-series data) 
and n=1 trials, were considered. The search results were imported into EndNote (version 9.3.2), 
duplicates were removed, and titles and abstracts were screened against
the inclusion criteria. Of the articles identified for full-text
screening, 10% were sampled, at random, for survey. Because \~30% of the
papers from the initial sample were not eligible for inclusion
(e.g., full-texts not in English, conference abstracts, qualitative
studies), an additional 5% random sample was taken. It was anticipated that 
30% of the additional 5% sample would also be ineligible. With this loss,
it was expected that 10% of the articles that remained after title and abstract 
screening would undergo survey. Figure 1 summarises the search process.

```{r prisma, fig.cap = "Flowchart of the article search and inclusion for the systematic search. The full-text of 137 articles was reviewed.", warning = FALSE, fig.height = 10, fig.width = 6}
knitr::include_graphics("figures/prisma-figure.pdf")
```

Articles were surveyed to determine: whether a missing data
statement was included, or if missing data were acknowledged in a
figure/table; and if not, whether articles shared their data. The articles
that included a missing data statement were surveyed to determine whether: 
(a) the type of missingness was reported; (b) missing data were explored; and
(c) if data were imputed. Two authors independently extracted the
data. Discrepancies were resolved by DNB. Findings are
reported as the proportion and 95% confidence interval (CI). Confidence
intervals were calculated using the Clopper-Pearson method for the
binomial distribution [@Clopper1934] via the *binom* package [@binom] in R [@rcore].

@@

The search returned a total of 3,863 articles. Of the 199 articles screened, `r n_screened` met the inclusion criteria, representing \~10% of the articles found (Figure 1). The proportion of articles that reported missing data was `r binom_ci_miss_reported$mean_pct` (95% CI; `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`). Of the articles that did not mention missing data (`r n_no_mention_missings`/`r n_screened`), only `r binom_ci_data_shared$mean_pct` (95% CI; `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`) shared their dataset, making it difficult to determine whether there was any missingness. While the absence of a missing data statement does not necessarily mean missing data are not being reported, it does suggest that this aspect of the analysis process receives little consideration. There were `r nrow(gs_football_missings_stated)` studies that provided a statement about missing data. A breakdown of the information extracted from these studies is provided in Table 1.

@@

A survey of `r n_screened` articles published in 2019 on football related topics
showed that only one in nine reported missing data. This result could suggest
that missing data are rarely considered by sport science researchers.
